---
title: "Tesla EV Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: cosmo
---

---

title: "Tesla EV Dashboard"
output:
flexdashboard::flex_dashboard:
orientation: columns
vertical_layout: fill
theme: cosmo
------------

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(flexdashboard)

theme_set(
  theme_minimal() +
    theme(
      plot.title = element_text(color = "#0B3C5D", size = 14, face = "bold"),
      axis.text = element_text(color = "#333333")
    )
)
```

```{r data}
# LOAD DATA
df <- read.csv("C:/Users/iyanu/Downloads/LEARNING DATA ANALYTICS/Learning R with Alex/Final Project/Files/Electric_Vehicle_Population_Data.csv")

# CLEAN
clean_df <- df %>%
  rename_all(~ gsub("\\.","_", .)) %>%
  mutate(
    Tesla = ifelse(Make == "TESLA", "TESLA", "OTHER"),
    Base_MSRP = as.numeric(Base_MSRP)
  )

# KPIs
total_vehicles <- nrow(clean_df)

tesla_vehicles <- clean_df %>%
  filter(Tesla == "TESLA") %>%
  nrow()

tesla_market_share <- round(tesla_vehicles / total_vehicles * 100, 1)

# MARKET SHARE
market_df <- data.frame(
  Group = c("TESLA", "OTHER"),
  Market_Share = c(tesla_market_share, 100 - tesla_market_share)
)

# MODELS SOLD
make_model_sold <- clean_df %>%
  filter(Tesla == "TESLA") %>%
  group_by(Model) %>%
  summarise(count_sold = n(), .groups = "drop") %>%
  arrange(desc(count_sold))

# PRICING DATA
df_tesla_prices <- read.csv("C:/Users/iyanu/Downloads/LEARNING DATA ANALYTICS/Learning R with Alex/Final Project/Files/Tesla_Current_Base_Prices.csv") %>%
  mutate(Model = toupper(Model))

Top_Tesla_Models_Priced <- make_model_sold %>%
  left_join(df_tesla_prices, by = "Model") %>%
  mutate(Estimated_revenue = count_sold * Base_Price_USD) %>%
  filter(!is.na(Estimated_revenue)) %>%
  group_by(Model) %>%
  slice_min(Base_Price_USD) %>%
  ungroup()

# TRENDS
tesla_sold_by_year <- clean_df %>%
  filter(Tesla == "TESLA") %>%
  group_by(Model_Year) %>%
  summarise(count = n(), .groups = "drop")

PHEV_vs_BEV <- clean_df %>%
  group_by(Model_Year, Electric_Vehicle_Type) %>%
  summarise(Count = n(), .groups = "drop")

# RANGE
Avg_Range_For_EVs <- clean_df %>%
  filter(Electric_Vehicle_Type == "Battery Electric Vehicle (BEV)",
         Electric_Range != 0,
         Base_MSRP != 0) %>%
  group_by(Tesla) %>%
  summarise(AVG_Range = mean(Electric_Range), .groups = "drop")

# UTILITIES
Utility_Count_for_Teslas <- clean_df %>%
  filter(Tesla == "TESLA") %>%
  group_by(Electric_Utility) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(desc(Count)) %>%
  head(5)
```

---

# 📊 Overview

## Column {data-width=350}

### Total Vehicles

```{r}
valueBox(total_vehicles, icon = "fa-car", color = "primary")
```

### Tesla Market Share

```{r}
valueBox(paste0(tesla_market_share, "%"), icon = "fa-chart-pie", color = "success")
```

### Tesla Vehicles

```{r}
valueBox(tesla_vehicles, icon = "fa-bolt", color = "info")
```

---

## Column

### Market Share

```{r}
ggplot(market_df, aes(x = "", y = Market_Share, fill = Group)) +
  geom_col(width = 1) +
  coord_polar("y") +
  scale_fill_manual(values = c("#1ABC9C", "#3498DB")) +
  theme_void()
```

---

# 🚗 Performance

## Column

### Models Sold

```{r}
ggplot(make_model_sold,
       aes(x = reorder(Model, count_sold), y = count_sold)) +
  geom_col(fill = "#3498DB") +
  coord_flip()
```

---

## Column

### Revenue per Model

```{r}
ggplot(Top_Tesla_Models_Priced,
       aes(x = reorder(Model, Estimated_revenue), y = Estimated_revenue)) +
  geom_col(fill = "#2ECC71") +
  coord_flip() +
  scale_y_continuous(labels = label_comma())
```

---

# 📈 Trends

## Column

### Tesla Sales Over Time

```{r}
ggplot(tesla_sold_by_year,
       aes(x = Model_Year, y = count)) +
  geom_line(color = "#3498DB") +
  geom_point(color = "#1ABC9C")
```

---

## Column

### PHEV vs BEV

```{r}
ggplot(PHEV_vs_BEV,
       aes(x = Model_Year, y = Count, color = Electric_Vehicle_Type)) +
  geom_line() +
  scale_color_manual(values = c("#3498DB", "#2ECC71"))
```

---

# ⚡ Insights

## Column

### Avg Range

```{r}
ggplot(Avg_Range_For_EVs,
       aes(x = Tesla, y = AVG_Range, fill = Tesla)) +
  geom_col() +
  scale_fill_manual(values = c("#2ECC71", "#3498DB"))
```

---

## Column

### Top Utilities

```{r}
ggplot(Utility_Count_for_Teslas,
       aes(x = reorder(Electric_Utility, Count), y = Count)) +
  geom_col(fill = "#1ABC9C") +
  coord_flip()
```

```